---
title: "Vandals"
author: "Ethan Marcano"
date: '2022-03-16'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy=TRUE, echo = TRUE)
library(tidyverse)
library(rpart)
library(rpart.plot)
library(randomForest)
library(caTools)
```

```{r datavandals}
vandals <- read_csv("Wikipedia.csv")
```

```{r vandal_count}
vandal_count <-vandals %>%
  select(everything()) %>%
  filter(Vandal == 1)
count(vandal_count)
```
+ There are 1815 cases of vandalism.
```{r wordschange}
mean(vandals$NumWordsAdded)
mean(vandals$NumWordsRemoved)
```

```{r correlations}
pairs(vandals)
cor(vandals)
```
+ Vandal is most negatively correlated to LoggedIn.

```{r vandalsplit}
set.seed(100)
spl = sample.split(vandals$Vandal, SplitRatio = 0.7)
vandalstrain = subset(vandals, spl==TRUE)
vandalstest = subset(vandals, spl==FALSE)

nrow(vandalstrain)/nrow(vandals)
nrow(vandalstest)/nrow(vandals)
```
### Question B)
```{r baseline_vandals_testing}
simple_vandal <- table(vandalstest$Vandal)
simple_vandal
simple_vandal[2]/sum(simple_vandal)
```
+ The baseline model is 47% accurate.

### Question C)

```{r vandal_validate}
set.seed(100)
spl <- sample.split(vandalstrain$Vandal, SplitRatio = 0.5)
vandals_validate_train <- subset(vandalstrain, spl == TRUE)
vandals_validate_test <- subset(vandalstrain, spl == FALSE)

vandal_CART1 <- rpart(Vandal ~ Minor + LoggedIn + HTTP + NumWordsAdded + NumWordsRemoved, method = "class", data = vandals_validate_train, minbucket = 5)
vandal_CART2 <- rpart(Vandal ~ Minor + LoggedIn + HTTP + NumWordsAdded + NumWordsRemoved, method = "class", data = vandals_validate_train, minbucket = 15)
vandal_CART3 <- rpart(Vandal ~ Minor + LoggedIn + HTTP + NumWordsAdded + NumWordsRemoved, method = "class", data = vandals_validate_train, minbucket = 25)

vandal_predict1 <- predict(vandal_CART1, newdata = vandals_validate_test, type = "class")
vandal_predict2 <- predict(vandal_CART2, newdata = vandals_validate_test, type = "class")
vandal_predict3 <- predict(vandal_CART3, newdata = vandals_validate_test, type = "class")

vandal_predict_table1 <- table(vandals_validate_test$Vandal, vandal_predict1)
vandal_predict_table2 <- table(vandals_validate_test$Vandal, vandal_predict2)
vandal_predict_table3 <- table(vandals_validate_test$Vandal, vandal_predict3)

vandal_predict_table1
sum(diag(vandal_predict_table1))/sum(vandal_predict_table1)

vandal_predict_table2
sum(diag(vandal_predict_table2))/sum(vandal_predict_table2)

vandal_predict_table3
sum(diag(vandal_predict_table3))/sum(vandal_predict_table3)
```

```{r vandal_CART}
vandal_CART <- rpart(Vandal ~ Minor + LoggedIn + HTTP + NumWordsAdded + NumWordsRemoved, method = "class", data = vandalstrain, minbucket = 25)
summary(vandal_CART)
prp(vandal_CART)
```

+ The relevant independent variables are LoggedIn, NumWordsAdded, and NumWordsRemoved. LoggedIn is the most relevant, with NumWordsRemoved branching off of NumWordsAdded.

#### ii)
```{r vandal_CART_predict}
vandal_CART_predict <- predict(vandal_CART, newdata = vandalstest, type = "class")
vandal_predict_CART_final <- table(vandalstest$Vandal, vandal_CART_predict)

vandal_predict_CART_final
sum(diag(vandal_predict_CART_final))/sum(vandal_predict_CART_final)
```
+ The model is 74.1% accurate.

### Question D)

```{r vandal_forest}
vandalstrain$Vandal <- as.factor(vandalstrain$Vandal)
vandalstest$Vandal <- as.factor(vandalstest$Vandal)

vandals_forest <- randomForest(Vandal ~ Minor + LoggedIn + HTTP + NumWordsAdded + NumWordsRemoved, data = vandalstrain, ntree = 200, nodesize = 15)
vandals_predict_forest <- predict(vandals_forest, newdata = vandalstest)
vandals_predict_final <- table(vandalstest$Vandal, vandals_predict_forest)

vandals_predict_final
sum(diag(vandals_predict_final))/sum(vandals_predict_final)
```
+ The model is slightly more accurate at ~75%.

### Question E)
#### i)
+ It can potentially be useful but for now it just gives hints as to who is committing vandalism, and the accuracy could be better with the introduction of more independent variables.

#### ii)
+ I would want to collect data on whether it was under an umbrella category (sports, politics, biology) or if the article in question was something that happened recently. Current events see more foot traffic and are heavily edited.

#### iii)
+ This model may not extend easily due to other pages often containing more contentious material, and thus subject to more vandalism as a whole.
